---
title: "Problem Set 1"
author: "Anaya Hall"
date: "January 27, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
```

***

## ARE 212 - Problem Set #1
Due Feb. 7, 2017  

***

### PART 1: Theory (Practice Only)

1-3. *come back to this!*

***

### PART 2: Applied - Carbon & Income

1. Download World Bank’s **World Development Indicators** (from Data Center): <http://databank.worldbank.org/data/reports.aspx?source=world-development-indicators>
Downloaded Jan. 27, 2018. 

2. Load data into R

```{r message=FALSE}

rm(list = ls())
mydir <- "~/projects/are212/pset1/"
datadir <- "wdi/"

# Bring in Data
setwd(mydir)
wdi_rawdata <- readr::read_csv(paste0(mydir,datadir,"wdi_carbon_income.csv"), na = ("..")
                            )
#Reshape data long to wide
wdi_df <- as.data.frame(wdi_rawdata)
wide <- reshape(wdi_df, idvar = c("Country Name", "Country Code"), timevar="Series Name", 
                drop = "Series Code", direction = "wide")

#Remove empty column of YEAR
wide <- wide[(-6)]

#Remove incomplete observations
wide <- na.omit(wide)

#Should be 194 countries. Check!
wdi_data <- rename(wide, CO2 = "2010 [YR2010].CO2 emissions (kt)", 
                  GDP =  "2010 [YR2010].GDP (constant 2010 US$)", 
                 POP = "2010 [YR2010].Population, total")

```

There are 194 countries with complete observations for the three series of interest. 

3. Calculate a table of showing the sample mean, standard deviation, minimum and maximum for each series:


```{r message=FALSE}
summary(wdi_data)
sapply(wdi_data, sd)
```


4. Create a histogram for CO2 and GDP (15 buckets).

```{r pressure}
hist(wdi_data$CO2, breaks=15)

hist(wdi_data$GDP)

```

```{r echo=FALSE}

ggplot(data=wdi_data, aes(x= GDP, y = CO2)) + geom_line() + ylab("CO2 Emissions") + xlab("GDP (USD 2010)") + labs(title="CO2 emissions against GDP (2010)")


```

