---
title: "ARE 212 Midterm"
author: "Anaya Hall"
date: "March 14, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```

# Question One: Replicate an important paper
## “Linkages among climate change, crop yields and Mexico - US cross-border migration"

Load OLS function

```{r OLS function, include=FALSE}
ols <- function(data, y_data, X_data, intercept = T, H0 = 0, two_tail = T, alpha = 0.05) {
  # Function setup ----
    # Require the 'dplyr' package
    require(dplyr)
    # Function to convert tibble, data.frame, or tbl_df to matrix
    to_matrix <- function(the_df, vars) {
      # Create a matrix from variables in var
      new_mat <- the_df %>%
        #Select the columns given in 'vars'
        select_(.dots = vars) %>%
        # Convert to matrix
        as.matrix()
      # Return 'new_mat'
      return(new_mat)
    }
  
  # Create dependent and independent variable matrices ----
    # y matrix
    y <- to_matrix (the_df = data, vars = y_data)
    # X matrix
    X <- to_matrix (the_df = data, vars = X_data)
      # If 'intercept' is TRUE, then add a column of ones
      if (intercept == T) {
      X <- cbind(1,X)
      colnames(X) <- c("intercept", X_data)
      }
 
  # Calculate b, y_hat, and residuals ----
    b <- solve(t(X) %*% X) %*% t(X) %*% y
    y_hat <- X %*% b
    e <- y - y_hat
  
  # Useful -----
    n <- nrow(X) # number of observations
    k <- ncol(X) # number of independent variables
    dof <- n - k # degrees of freedom
    i <- rep(1,n) # column of ones for demeaning matrix
    A <- diag(i) - (1 / n) * i %*% t(i) # demeaning matrix
    y_star <- A %*% y # for SST
    X_star <- A %*% X # for SSM
    SST <- drop(t(y_star) %*% y_star)
    SSM <- drop(t(b) %*% t(X_star) %*% X_star %*% b)
    SSR <- drop(t(e) %*% e)
  
  # Measures of fit and estimated variance ----
    R2uc <- drop((t(y_hat) %*% y_hat)/(t(y) %*% y)) # Uncentered R^2
    R2 <- 1 - SSR/SST # Uncentered R^2
    R2adj <- 1 - (n-1)/dof * (1 - R2) # Adjusted R^2
    AIC <- log(SSR/n) + 2*k/n # AIC
    SIC <- log(SSR/n) + k/n*log(n) # SIC
    s2 <- SSR/dof # s^2
  
  # Measures of fit table ----
    mof_table_df <- data.frame(R2uc, R2, R2adj, SIC, AIC, SSR, s2)
    mof_table_col_names <- c("$R^2_\\text{uc}$", "$R^2$",
                             "$R^2_\\text{adj}$",
                             "SIC", "AIC", "SSR", "$s^2$")
    mof_table <-  mof_table_df %>% knitr::kable(
      row.names = F,
      col.names = mof_table_col_names,
      format.args = list(scientific = F, digits = 4),
      booktabs = T,
      escape = F
    )
  
  # t-test----
    # Standard error
    se <- as.vector(sqrt(s2 * diag(solve(t(X) %*% X))))
    # Vector of _t_ statistics
    t_stats <- (b - H0) / se
    # Calculate the p-values
    if (two_tail == T) {
    p_values <- pt(q = abs(t_stats), df = dof, lower.tail = F) * 2
    } else {
      p_values <- pt(q = abs(t_stats), df = dof, lower.tail = F)
    }
    # Do we (fail to) reject?
    reject <- ifelse(p_values < alpha, reject <- "Reject", reject <- "Fail to Reject")
    
    # Nice table (data.frame) of results
    ttest_df <- data.frame(
      # The rows have the coef. names
      effect = rownames(b),
      # Estimated coefficients
      coef = as.vector(b) %>% round(3),
      # Standard errors
      std_error = as.vector(se) %>% round(3),
      # t statistics
      t_stat = as.vector(t_stats) %>% round(3),
      # p-values
      p_value = as.vector(p_values) %>% round(4),
      # reject null?
      significance = as.character(reject)
      )
  
    ttest_table <-  ttest_df %>% knitr::kable(
      booktabs = T,
      format.args = list(scientific = F),
      escape = F
    )

  # Data frame for exporting for y, y_hat, X, and e vectors ----
    export_df <- data.frame(y, y_hat, e, X) %>% tbl_df()
    colnames(export_df) <- c("y","y_hat","e",colnames(X))
  
  # Return ----
    return(list(n=n, dof=dof, b=b, vars=export_df, R2uc=R2uc,R2=R2,
                R2adj=R2adj, AIC=AIC, SIC=SIC, s2=s2, SST=SST, SSR=SSR,
                mof_table=mof_table, ttest=ttest_table))
}
```

```{r}
# Function to turn given data into matrix for use in OLS function
to_matrix <- function(the_df, vars) {
  # Create a matrix from variables in var
  new_mat <- the_df %>%
    # Select the columns given in 'vars'
    select_(.dots = vars) %>%
    # Convert to matrix
    as.matrix()
  # Return 'new_mat'
  return(new_mat)
}

# Function for OLS coefficient estimates and measures of fit
b_ols <- function(data, y_data, X_data, intercept=TRUE) {

  require(dplyr)  
    # y matrix
    y <- to_matrix (the_df = data, vars = y_data)
    # X matrix
    X <- to_matrix (the_df = data, vars = X_data)
      # If 'intercept' is TRUE, then add a column of ones
      if (intercept == T) {
        X <- cbind(1,X)
        colnames(X) <- c("intercept", X_data)
      }

    # Calculate beta hat ------
    b <- solve( t(X) %*% X ) %*% t(X) %*% y
    # Change the name of 'ones' to 'intercept'
    if(intercept == T){
        rownames(b) <- c("intercept", X_data) }
    else
        rownames(b) <- c(X_data)
    
    y_hat <- X %*% b
    e <- y - y_hat
  
    # Useful transformations -----
    n <- nrow(X) # number of observations
    k <- ncol(X) # number of independent variables
    dof <- n - k # degrees of freedom
    i <- rep(1,n) # column of ones for demeaning matrix
    A <- diag(i) - (1 / n) * i %*% t(i) # demeaning matrix
    y_star <- A %*% y # for SST
    X_star <- A %*% X # for SSM
    SST <- drop(t(y_star) %*% y_star)
    SSM <- drop(t(b) %*% t(X_star) %*% X_star %*% b)
    SSR <- drop(t(e) %*% e)
  
    # Measures of fit and estimated variance ----
    R2uc <- drop((t(y_hat) %*% y_hat)/(t(y) %*% y)) # Uncentered R^2
    R2 <- 1 - SSR/SST # Uncentered R^2
    R2adj <- 1 - (n-1)/dof * (1 - R2) # Adjusted R^2
    AIC <- log(SSR/n) + 2*k/n # AIC
    SIC <- log(SSR/n) + k/n*log(n) # SIC
    s2 <- SSR/dof # s^2
  
    results <- data.frame(
      # The rows have the coef. names
      x_var = rownames(b),
      # Estimated coefficients
      coef = as.vector(b) %>% round(3)
    )

      # Return beta_hat & adjusted r2
    #return(R2adj)
    return(results)
}

```


```{r}

b_ols(data = auto, y_data = "price", X_data = c("mpg", "headroom"))
```


Load & clean data

1. Estimate model (1) via OLS by regressing emigration rate on log of yields and a time period fixed effect.Report coefficient on yield and adjusted $R^2$. Does this match the results in the first column of table #1?

2. Estimate model (1) again via fixed effects and FWT. Report coefficient on yield and adjusted $R^2$. Does this match the results in the third column of table #1?

3. Repeat step 1 without the the fixed effects. Report coefficient on yield and adjusted $R^2$. Do the results look different from what you estimated before? From what is in the paper?

4. Repeat step 2 without the the fixed effects. Report coefficient on yield and adjusted $R^2$. Do the results look different from what you estimated before? From what is in the paper?

5. What happened here? What are the consequences?


```{r}


```

# Question Two: Normality of OLS

Model:
$y_i$ = $\beta_o$ + $\beta_1$$x_{1i}$ + $\beta_2$$x_{2i}$ + $\epsilon_i$

Truth:
$\beta_0$ = 3 , $\beta_1$ = 1, $\beta_2$ = -2


###Load functions for use in simulation

Generate data function (given a sample size, n)
```{r gen_data function}

gen_data <- function(sample_size) {
  # Create data.frame with random x and error
  data_df <- data.frame(
    x1 = rnorm(sample_size),
    x2 = rnorm(sample_size),
    e = rnorm(sample_size))
  # Calculate y = 3 + 1 x1 - 2 x2 + e; drop 'e'
  data_df %<>% mutate(y = 3 + 1 * x1 - 2 * x2 + e) %>%
    select(-e)
  # Return data_df
  return(data_df)
  
}
```

Run a single simulation of OLS
```{r one_sim function}
one_sim <- function(sample_size) {
  # Estimate via OLS
  ols_est <- ols(data = gen_data(sample_size),
    y_data = "y", X_data = c("x1", "x2"))
  # Grab the estimated coefficient on x
  # (the second element of 'coef')
  b2 <- ols_est %$% coef[3]
  # Grab the second p-value
  # (the first p-value is for the intercept)
  # Return a data.frame with b1 and p_value
  return(data.frame(b1, p_value))
}
```
