---
title: "ARE 212 Midterm"
author: "Anaya Hall"
date: "March 14, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(readr)
library(magrittr)
library(pacman)
p_load(dplyr, lfe, magrittr, readr, ggplot2)
```

# Question One: “Linkages among climate change, crop yields and Mexico - US cross-border migration"

Load OLS functions

```{r OLS functions}
# Function to turn given data into matrix for use in OLS function
to_matrix <- function(the_df, vars) {
  # Create a matrix from variables in var
  new_mat <- the_df %>%
    # Select the columns given in 'vars'
    select_(.dots = vars) %>%
    # Convert to matrix
    as.matrix()
  # Return 'new_mat'
  return(new_mat)
}

# Function for OLS coefficient estimates and measures of fit
b_ols <- function(data, y_data, X_data, intercept=TRUE) {

  require(dplyr)  
    # y matrix
    y <- to_matrix (the_df = data, vars = y_data)
    # X matrix
    X <- to_matrix (the_df = data, vars = X_data)
      # If 'intercept' is TRUE, then add a column of ones
      if (intercept == T) {
        X <- cbind(1,X)
        colnames(X) <- c("intercept", X_data)
      }

    # Calculate beta hat ------
    b <- solve( t(X) %*% X ) %*% t(X) %*% y
    # Change the name of 'ones' to 'intercept'
    if(intercept == T){
        rownames(b) <- c("intercept", X_data) }
    else
        rownames(b) <- c(X_data)
    
    y_hat <- X %*% b
    e <- y - y_hat
  
    # Useful transformations -----
    n <- nrow(X) # number of observations
    k <- ncol(X) # number of independent variables
    dof <- n - k # degrees of freedom
    i <- rep(1,n) # column of ones for demeaning matrix
    A <- diag(i) - (1 / n) * i %*% t(i) # demeaning matrix
    y_star <- A %*% y # for SST
    X_star <- A %*% X # for SSM
    SST <- drop(t(y_star) %*% y_star)
    SSM <- drop(t(b) %*% t(X_star) %*% X_star %*% b)
    SSR <- drop(t(e) %*% e)
  
    # Measures of fit and estimated variance ----
    R2uc <- drop((t(y_hat) %*% y_hat)/(t(y) %*% y)) # Uncentered R^2
    R2 <- 1 - SSR/SST # Uncentered R^2
    R2adj <- 1 - (n-1)/dof * (1 - R2) # Adjusted R^2
    AIC <- log(SSR/n) + 2*k/n # AIC
    SIC <- log(SSR/n) + k/n*log(n) # SIC
    s2 <- SSR/dof # s^2
  
    results <- data.frame(
      # The rows have the coef. names
      x_var = rownames(b),
      # Estimated coefficients
      coef = as.vector(b) %>% round(3)
    )

    res <<- e
    adjr2 <<- R2adj
    
    # Return beta_hat & adjusted r2
    #print(paste("Adj R2: ", R2adj))
    return(results)
}

```

Load Data (cleaned and prepared in Excel)
```{r load_data, echo=FALSE}
feng_data <- read_csv("~/projects/are212/midterm/feng_data.csv")
```

Inspect data
```{r}
summary(feng_data)
```

1. **Estimate model (1) via OLS by regressing emigration rate on log of yields and a time period fixed effect. Report coefficient on yield and adjusted $R^2$. Does this match the results in the first column of table #1?**

```{r cornols, echo=FALSE}
b_ols(data = feng_data, y_data = "remi", X_data = c("cornyield", "y1995"))
```
Adjusted $R^2$: 0.321155954568775


*These don't really match the results in the paper*

```{r cornwheatols, echo=FALSE}
b_ols(data = feng_data, y_data = "remi", X_data = c("cornwheatyield", "y1995"))
```
Adjusted $R^2$: 0.32147846639082"


*These don't really match the results in the paper*

2. **Estimate model (1) again via fixed effects and FWT. Report coefficient on yield and adjusted $R^2$. Does this match the results in the third column of table #1?**

```{r resid_OLS}
resid_ols <- function(data, y_var, X_vars, intercept = T) {
  # Require the 'dplyr' package
  require(dplyr)
  # Create the y matrix
  y <- to_matrix(the_df = data, vars = y_var)
  # Create the X matrix
  X <- to_matrix(the_df = data, vars = X_vars)
  # If 'intercept' is TRUE, then add a column of ones
  if (intercept == T) {
    # Bind a column of ones to X
    X <- cbind(1, X)
    # Name the new column "intercept"
    colnames(X) <- c("intercept", X_vars)
  }
  # Calculate the sample size, n
  n <- nrow(X)
  # Calculate the residuals
  resids <- (diag(n) - X %*% solve(t(X) %*% X) %*% t(X)) %*% y
  # Return 'resids'
  return(resids)
}
```


```{r FWT_corn}
step1_resid <- resid_ols(feng_data, "remi", "cornyield", F)

feng_data %<>% mutate(ones = 1)
# Our two regressions
step2a_resid <- resid_ols(feng_data, "ones", "cornyield", F)
step2b_resid <- resid_ols(feng_data, "y1995", "cornyield", F)


df_fwt <- data.frame(
  remi_resid  = as.vector(step1_resid),
  i_resid      = as.vector(step2a_resid),
  fe_resid = as.vector(step2b_resid)
  ) %>% tbl_df()
# The final regression
b_ols(df_fwt, "remi_resid", c("i_resid", "fe_resid"), F)

```
Adjusted $R^2$: 0.5573145

```{r FWT_cornwheat}
step1_resid <- resid_ols(feng_data, "remi", "cornwheatyield", F)

feng_data %<>% mutate(ones = 1)
# Our two regressions
step2a_resid <- resid_ols(feng_data, "ones", "cornwheatyield", F)
step2b_resid <- resid_ols(feng_data, "y1995", "cornwheatyield", F)


df_fwt <- data.frame(
  remi_resid  = as.vector(step1_resid),
  i_resid      = as.vector(step2a_resid),
  fe_resid = as.vector(step2b_resid)
  ) %>% tbl_df()
# The final regression
b_ols(df_fwt, "remi_resid", c("i_resid", "fe_resid"), F)

```
Adjusted $R^2$: 0.5556311

*These both match the output from my first regressions, but they do not match the results in the paper*

3. **Repeat step 1 without the the fixed effects. Report coefficient on yield and adjusted $R^2$. Do the results look different from what you estimated before? From what is in the paper?**

```{r}
b_ols(data = feng_data, y_data = "remi", X_data = c("cornyield"))

```
Adjusted $R^2$: -0.004523002


```{r}
b_ols(data = feng_data, y_data = "remi", X_data = c("cornwheatyield"))

```
Adjusted $R^2$: -0.002158803


4. **Repeat step 2 without the the fixed effects. Report coefficient on yield and adjusted $R^2$. Do the results look different from what you estimated before? From what is in the paper?**

```{r FWT_corn_nofe}
step1_resid <- resid_ols(feng_data, "remi", "cornyield", F)

feng_data %<>% mutate(ones = 1)
# Our two regressions
step2a_resid <- resid_ols(feng_data, "ones", "cornyield", F)


df_fwt <- data.frame(
  remi_resid  = as.vector(step1_resid),
  i_resid      = as.vector(step2a_resid)
  ) %>% tbl_df()
# The final regression
b_ols(df_fwt, "remi_resid", c("i_resid"), F)

```
Adjusted $R^2$:  0.3447634

```{r FWT_cornwheat_nofe}
step1_resid <- resid_ols(feng_data, "remi", "cornwheatyield", F)

feng_data %<>% mutate(ones = 1)
# Our two regressions
step2a_resid <- resid_ols(feng_data, "ones", "cornwheatyield", F)

df_fwt <- data.frame(
  remi_resid  = as.vector(step1_resid),
  i_resid      = as.vector(step2a_resid)
  ) %>% tbl_df()
# The final regression
b_ols(df_fwt, "remi_resid", c("i_resid"), F)

```

Adjusted $R^2$: 0.3435078

*The results still do not match the output...*



5. **What happened here? What are the consequences?**

*Our regressions here do not replicate the results from the Feng et al paper. Although we included period fixed effects, we did not control for any state-level fixed effects. This could be hugely influential on emigration ratio, considering that local policies and other cultural on-goings could be arguably more impactful on citizen's immigration choices. We therefore could be missing a lot of the explanatory power of emigration in Mexican states.*


# Question Two: Normality of OLS

Model:
$y_i$ = $\beta_o$ + $\beta_1$$x_{1i}$ + $\beta_2$$x_{2i}$ + $\epsilon_i$

Truth:
$\beta_0$ = 3 , $\beta_1$ = 1, $\beta_2$ = -2

###Load functions for use in simulation

Generate data function (given a sample size, n)
```{r slimols}

# Skinny version of OLS function without all the measures of fit
slim_ols <- function(data, y_data, X_data, intercept=TRUE) {

  require(dplyr)  
    # y matrix
    y <- to_matrix (the_df = data, vars = y_data)
    # X matrix
    X <- to_matrix (the_df = data, vars = X_data)
      # If 'intercept' is TRUE, then add a column of ones
      if (intercept == T) {
        X <- cbind(1,X)
        colnames(X) <- c("intercept", X_data)
      }

    # Calculate beta hat ------
    b <- solve( t(X) %*% X ) %*% t(X) %*% y
    # Change the name of 'ones' to 'intercept'
    if(intercept == T){
        rownames(b) <- c("intercept", X_data) }
    else
        rownames(b) <- c(X_data)
 
    results <- data.frame(
      # The rows have the coef. names
      x_var = rownames(b),
      # Estimated coefficients
      coef = as.vector(b) %>% round(3)
    )
    
    return(results)   
}

```


```{r gen_data function}
#ADD SWEEP TO FIX COVARIANCE OF X1 & X2 (Should be 0, but isn't right now) 
gen_data <- function(sample_size) {
  # Create data.frame with random x and error
  data_df <- data.frame(
    X1 = rnorm(n = sample_size, sd = 5),
    X2 = rnorm(n = sample_size, sd = 5),
    # X <- rnorm(n = sample_size, sd =5) %>% matrix(sample_size, 2),
    e = rnorm(sample_size, sd = 5),
    eta = runif(sample_size, -8.66, 8.66))
  # Calculate y = 3 + 1 x1 - 2 x2 + e; drop 'e'
  data_df %<>% mutate(y_a = 3 + 1 * X1 - 2 * X2 + e, 
                      y_b = 3 + 1 * X1 - 2 * X2 + eta) %>% #Populaton
    select(-e, -eta)
  # Return data_df
  return(data_df)
  
}

```

Function for a single simulation of OLS
```{r one_sim function}
one_sim <- function(sample_size, depvar) {
  # Estimate via OLS
  ols_est <- slim_ols(data = gen_data(sample_size),
    y_data = depvar, X_data = c("X1", "X2"))
  # Grab the estimated coefficient on x
  # (the second element of 'coef')
  b2 <- ols_est %$% coef[3]
  # Return a data.frame with b2
  return(data.frame(b2))
}
```


Function for multiple simulations of OLS
```{r}
ols_sim <- function(depvar, n_sims = 1e4, sample_size, seed = 22092008) {
  # require parallel
  require(parallel)
  # Set the seed
  set.seed(seed)
  # Run one_sim n_sims times; convert results to data.frame
  sim_df <- replicate(
    n = n_sims,
    expr = one_sim(sample_size, depvar),
    simplify = F
    ) %>% bind_rows()
  
   # TRY TO PARALLELIZE TO SPEED UP! ---- get error with bind rows though...
   # sim_df <- mclapply(
   #  X = rep(x = sample_size, times = n_sims),
   #  FUN = one_sim,
   #  # Specify that we want 4 cores
   #  mc.cores = 4
   #  ) %>% bind_rows()
  # Return sim_df
  return(sim_df)
}
```

For each part, repeat for sample sizes: n=[10, 100, 1000, 10000, 20000] and run 1e4 simulations


###Part A: 
Regress $y^a$ on intercept, $x_1$ and $x_2$. Record $\beta_2$ 
```{r ya_sims}
N <- c(10, 100, 1000, 10000, 20000)

# sim_A <- matrix("list", 5)
# for (n in N){
#   print(n)
#   sim_A[[n]] <- ols_sim(depvar = "y_a", n_sims = 1e4, sample_size = n)
# 
# }

# Run for sample sizes: n=[10, 100, 1000, 10000, 20000]
# Run ols_sim for sample size of 10
sima10 <- ols_sim(depvar = "y_a", n_sims = 1e4, sample_size = 10)
# Run ols_sim for sample size of 100
sima100 <- ols_sim(depvar = "y_a", n_sims = 1e4, sample_size = 100)
# Run ols_sim for sample size of 1000
sima1000 <- ols_sim(depvar = "y_a", n_sims = 1e4, sample_size = 1000)
# # Run ols_sim for sample size of 10000
sima10000 <- ols_sim(depvar = "y_a", n_sims = 1e4, sample_size = 10000)
# # Run ols_sim for sample size of 20000
sima20000 <- ols_sim(depvar = "y_a", n_sims = 1e4, sample_size = 20000)

```

Plot histograms

```{r a_plots}
ggplot(sima10, aes(x=b2)) + geom_density() + stat_function(fun=dnorm, args= list(mean= -2), color="blue") + labs(title="10 Iterations") 
```
```{r}
ggplot(sima100, aes(x=b2)) + geom_density() + stat_function(fun=dnorm, args= list(mean= -2), color="blue") + labs(title="100 Iterations")
```
```{r}
ggplot(sima1000, aes(x=b2)) + geom_density() + stat_function(fun=dnorm, args= list(mean= -2), color="blue") + labs(title="1000 Iterations")
```
```{r}
ggplot(sima10000, aes(x=b2)) + geom_density() + stat_function(fun=dnorm, args= list(mean= -2), color="blue") + labs(title="10000 Iterations")
```
```{r}
ggplot(sima20000, aes(x=b2)) + geom_density() + stat_function(fun=dnorm, args= list(mean= -2), color="blue") + labs(title="20000 Iterations")
```


###Part B:
Regress $y^b$ on $x_1$ and $x_2$. Record $\beta_2$ 

```{r yb_sims}
# Run for sample sizes: n=[10, 100, 1000, 10000, 20000]
# Run ols_sim for sample size of 10
simb10 <- ols_sim(depvar = "y_b", n_sims = 1e4, sample_size = 10)
# Run ols_sim for sample size of 100
simb100 <- ols_sim(depvar = "y_b", n_sims = 1e4, sample_size = 100)
# # Run ols_sim for sample size of 1000
simb1000 <- ols_sim(depvar = "y_b", n_sims = 1e4, sample_size = 1000)
# # Run ols_sim for sample size of 10000
simb10000 <- ols_sim(depvar = "y_b", n_sims = 1e4, sample_size = 10000)
# # Run ols_sim for sample size of 20000
simb20000 <- ols_sim(depvar = "y_b", n_sims = 1e4, sample_size = 20000)
```

Plot histogram

``` {r}
ggplot(simb10, aes(x=b2)) + geom_density() + stat_function(fun=dnorm, args= list(mean= -2), color="blue") + labs(title="10 Iterations")
```
```{r}
ggplot(simb100, aes(x=b2)) + geom_density() + stat_function(fun=dnorm, args= list(mean= -2), color="blue") + labs(title="100 Iterations")
```
```{r}
ggplot(simb1000, aes(x=b2)) + geom_density() + stat_function(fun=dnorm, args= list(mean= -2), color="blue") + labs(title="1000 Iterations")
```
```{r}
ggplot(simb10000, aes(x=b2)) + geom_density() + stat_function(fun=dnorm, args= list(mean= -2), color="blue") + labs(title="10000 Iterations")
```
```{r}
ggplot(simb20000, aes(x=b2)) + geom_density() + stat_function(fun=dnorm, args= list(mean= -2), color="blue") + labs(title="N = 20000")

```

**Thanks to Ed for providing A LOT of this code in section!**